---
title: "iModMix -integrative Modules for Multi-omics data"
author: "Isis Narvaez-Bandera"
date: "`r Sys.Date()`"
package: "`r packageVersion('iModMix')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: united
    highlight: tango
fig_width: 7
vignette: >
    %\VignetteIndexEntry{Publication-ready integration omics data including unidentified metabolites}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    %\usepackage[utf8]{inputenc}
---

# Introducción

`iModMix` is a novel network-based approach that integrates multi-omics data, including metabolomics, proteomics, and transcriptomics data, into a unified network to unveil associations across various layers of omics data and reveal significant associations with phenotypes of interest.

iModMix can incorporate both identified and unidentified metabolites, addressing a key limitation of existing methodologies.

```{r, echo=FALSE}

  suppressWarnings(library(knitr))

  opts_chunk$set(tidy = FALSE, message = FALSE, warning = FALSE)

```

# Installation

## Download the package from Bioconductor

```{r getPackage, eval=FALSE}

  if (!requireNamespace('BiocManager', quietly = TRUE))
    install.packages('BiocManager')

  BiocManager::install('iModMix')

```

Note: to install development version:

```{r getPackageDevel, eval=FALSE}

  devtools::install_github('biodatalab/iModMix')

```

## Load the package into R session

```{r Load, message=FALSE}

  library(iModMix)

```

# Example to integrate 2 omics datasets

## Generate modules

### Load expresion Data 1 and metadata from clear cell renal cell carcinoma

This is a basic example which shows you how to analyze two datasets through iModMix package: We used 24 normal and 52 tumor clear cell renal cell carcinoma (ccRCC) samples (Golkaram et al. 2022; Tang et al. 2023; Benedetti et al. 2023) as a case study. It contained 23001 genes from RNA-seq and 904 identified metabolites from untargeted metabolomics.

Input data: 

* The expression datasets would be a matrix with genes as rows and samples as columns
* The meta data should be a matrix with the first column named 'Sample' (note the capitalized 'S'). The second and others column names can be arbitrary, i.e. "disease" or "Genotype"

```{r}
# Load the package
library(iModMix)

# Get the path to the expression data file
path_Metab_exp <- system.file("Example_data/ccRCC4_Data/Metab_exp.rds", package = "iModMix")

# Load the expression data
Data_exp1 <- readRDS(path_Metab_exp)

# Check the expression data
Data_exp1[1:5, 1:5]

# Get the path to the Metadata file
path_Metadata <- system.file("Example_data/ccRCC4_Data/Metadata.rds", package = "iModMix")

# Load the Metadata
Metadata <- readRDS(path_Metadata)

# Check the Metadata
head(Metadata)

```

### Perform preprocessing data

The iModMix pipeline is expecting an input matrix of RNA Sequence counts. Usually we need to rotate (transpose) the input data so rows = treatments and columns = features.

```{r}
# Preprocess de data in the format require for iModMix
load_data1 <- load_data(Data_exp1)
load_data1[1:5, 1:5]
```

### Perform Partial correlation, network construction and module eigengene calculation

Partial correlation is usually the slowest step. Partial correlation is a method of analyzing the relationship between two variables when other variables are present. Graphical Lasso (Glasso) is used to estimate the partial correlation and captures only direct associations. Below is a preview of the sparse partial correlation of the first five features. 

```{r}
# Perform Partial correlation 
parcorData1 <- partial_cors(load_data1, rho = 0.25)
parcorData1[1:5, 1:5]
```
Hierarchical clustering is used to identify common neighbors between the features. Calculations are determined using the topographical overlap matrix (TOM) and based on the sparse partial correlations. Hierarchical clustering is visualized as a dendrogram. 

The dendogram plot is publication-ready and display a dendrogram of features and modules. 
Axes: The vertical axis (y-axis) represents the dissimilarity between features, while the horizontal axis (x-axis) shows the modules. 
Branches: Each line in the dendrogram represents a feature. Features that are closer in the hierarchy (i.e., joined at a lower height in the dendrogram) have more similar expression profiles.


```{r}
# Perform hierarchical clustering 
hcData1 <- hierarchical_cluster(parcor_mat = parcorData1, tom = TRUE, min_module_size = 10)

hcClu = hcData1$hclustTree
hcMod = as.matrix(hcData1$dynamicMods_numeric)
WGCNA::plotDendroAndColors(dendro = hcClu,
                                 colors = hcMod,
                                 dendroLabels = FALSE,
                                 hang = 0.03,
                                 addGuide = TRUE,
                                 guideHang = 0.05,
                                 groupLabels = "Modules",
                                 main = "Feature dendrogram and module assignments")
```

Let's see the cluster assignments. Hierarchical clustering generates multiple modules (clusters) to which each feature is assigned. The table below details the following columns:

* **Feature**: Feature ID.
* **cluster**: The module where the feature is assigned 
* **col**: The color used on the hierarchical clustering dendrogram.

```{r}
# Perform cluster assignments
cluster_assig_Data1 <- cluster_assignments(as.data.frame(hcData1$cluster_assignments))
head(cluster_assig_Data1)
```

The 904 features were assigned into 34 modules. The first principal component (PC1) is calculated for each module, referred to as an eigenfeature. Eigenfeatures are useful for

* Relating the modules to the phenotypes. 
* Obtaining the correlation between omics datasets (integration). 

```{r}
# Obtain Eigenfeatures
eigenData1 <- Eigengenes(load_data1, hcData1$cluster_assignments[,3])
eigengenes_Data1 <- eigenData1$module_eigenmetab_Me
eigengenes_Data1[1:5, 1:5]
```

### Load expresion Data 2 and perform partial correlation, network construction and module eigengene calculation

```{r}
# Get the path to the expression data file
path_RNA_exp <- system.file("Example_data/ccRCC4_Data/RNA_exp.rds", package = "iModMix")

# Load the expression data
Data_exp2 <- readRDS(path_RNA_exp)
Data_exp2 <- Data_exp2[1:1000,]

# Check the expression data
Data_exp2[1:5, 1:5]

#Preprocess de data in the format require for iModMix
load_data2 <- load_data(Data_exp2)
```

Calculate correlated matrix  

```{r}
# Partial correlation
parcorData2 <- partial_cors(load_data2, rho = 0.25) 
parcorData2[1:5, 1:5]
```

Calculate hierarchical clustering 

```{r}
# Perform hierarchical clustering 
hcData2 <- hierarchical_cluster(parcor_mat = parcorData2, tom = TRUE, min_module_size = 10) 

# Access the hierarchical clustering tree and module assignments
hcClu2 <- hcData2$hclustTree
hcMod2 <- as.matrix(hcData2$dynamicMods_numeric)

# Plot the dendrogram
WGCNA::plotDendroAndColors(dendro = hcClu2,
                          colors = hcMod2,
                          dendroLabels = FALSE,
                          hang = 0.03,
                          addGuide = TRUE,
                          guideHang = 0.05,
                          groupLabels = "Modules",
                          main = "Feature dendrogram and module assignments")

```

Perform network construction and module eigengene calculation

```{r}
# Perform cluster assignment
cluster_assig_Data2 <- cluster_assignments(as.data.frame(hcData2$cluster_assignments))
eigenData2 <- Eigengenes(load_data2, hcData2$cluster_assignments[,3])
eigengenes_Data2 <- eigenData2$module_eigenmetab_Me
eigengenes_Data2[1:5, 1:5]
```
## Relate modules to phenotype

Statistical analysis by Students t-test compares phenotypes chosen as variable of interest. The eigenfeatures of each module, determined previously, are used as predictors. The user can also specify a significance threshold for the p-value, with the default set to 0.05. It returns a data frame with the following columns 

Here, TN variable (Tumor, Normal) is selected as the phenotype variable of interest 

Boxplots are automatically generated at the bottom for significant eigenfeatures, with dots marking outliers and a legend describing the compared phenotype.


```{r}
# Perform Classification
Classification_Data <- perform_classification(eigengene_data = eigengenes_Data1,
                                              metadata = Metadata,
                                              phenotype_variable = "TN",
                                              significance_threshold = 0.09)

Classification_Data$result[1:10,]

# Plot BoxPlot
selected_variable <- "TN"
levels <- unique(Metadata[[selected_variable]])
class_label <- paste(levels, collapse = " vs ")
plot <- Classification_Data$plots[[1]]
        plot <- plot +
          ggplot2::labs(title = class_label, fill = as.factor(levels),
                        x = "Variables",
                        y = "Class") +
          ggplot2::theme(
            axis.text.x = ggplot2::element_text(angle = 90, hjust = 1)
          )
plot
```

## Multi-omics analysis: Integration of Data 1 and Data 2

This step perform the integration of the datasets in the analysis. Is necessary list the eigenfuetues obtained from the previus steps. 

```{r}
# Summarize into list the diffrent egenfuatures 
eigengenes_list <- list(eigengenes_Data1, eigengenes_Data2)
cluster_list <- list(hcData1$cluster_assignments, hcData2$cluster_assignments)
Integration_Data1_Data2 <- Modules_correlation(eigengenes_list, cluster_list, threshold = 0.6)
TopCorrelations <-Integration_Data1_Data2[["Top_cor_Prot_metab"]]
head(TopCorrelations)

# Plot correlation
Correlation_Plot <- Integration_Data1_Data2$Correlation_Plot
hist(Correlation_Plot[[1]], main = "Correlation: Data 1 / Data 2 ")

#Plot Network
nodes <- as.data.frame(Integration_Data1_Data2$nodes)
edges <- as.data.frame(Integration_Data1_Data2$edges)
n <- Integration_Data1_Data2$n

shapes <- c("diamond", "triangle", "dot")
colors <- c("orange", "darkgreen", "darkblue")

network <- visNetwork::visNetwork(nodes = nodes, edges = edges, width = "100%", height = "800px")
network <- visNetwork::visLegend(network, useGroups = FALSE, 
                                 addNodes = data.frame(label = paste0("Data", 1:n, " Modules"),
                                 shape = shapes[1:n], color = colors[1:n]),
                                 addEdges = data.frame(label = "Correlation", shape = "line", length = 200, color = "darkgreen"))
network <- visNetwork::visInteraction(network, navigationButtons = TRUE)
network
      
```

